Modal Verification Patterns for Systems

### 2025-06-16 paper.tex
+

\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}. \textsf{Perennial}~\cite{chajed2019verifying} extends the weakest precondition of the \textsf{Iris} separation logic~\cite{jung2018iris,krebbers2017essence} to allow asserting facts about disk state at crashes (the \emph{crash condition}) and apply this extension to a journalling filesystem \textsf{GoJournal}~\cite{chajed2021gojournal}. Unlike \textsf{GoJournal}, \textsf{ModFs} relies on \textsf{CoW} semantics for consistency, eliminating the need for recovery logic.
Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
\paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory. To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cow_index}, from which a path is created for a stack pointer reference to the allocated memory. To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime. 
\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the condition where there is no \emph{reachability} to the reference pointing to the allocated memory reference, and it is not \emph{shared}. Once the requirements are satisfied then the memory location can be reclaimed \emph{soundly} -- i.e., without violating any access through a reference to it.
\subsection{Conclusion \& Contributions}

>>>>

\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}In-place updating filesystems with journalling components have been specified and verified with an extended Hoare Logic to specify not only the pre/post conditions of the operations successful operations in the filesystem, but also the crash condition~\cite{chen-crash-16,chen2017verifying}, with later work extending this for concurrent filesystem implementations~\cite{chajed2021gojournal}. \textsf{Perennial}~\cite{chajed2019verifying} extends the weakest precondition of the \textsf{Iris} separation logic~\cite{jung2018iris,krebbers2017essence} to allow asserting facts about disk state at crashes (the \emph{crash condition}) and apply this extension to a journalling filesystem \textsf{GoJournal}~\cite{chajed2021gojournal}. Unlike \textsf{GoJournal}, \textsf{ModFs} relies on \textsf{CoW} semantics for consistency, eliminating the need for recovery logic.
Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cow_index_tree_update}, therefore, considering \textsc{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
\paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory. To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cow_index}, from which a path is created for a stack pointer reference to the allocated memory. To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime. 
\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the conditions where the unique is not \emph{shared}, and there is no \emph{reachability} from the root to any the references pointing to the allocated memory. Once the requirements are satisfied then the memory location can be reclaimed \emph{soundly} -- i.e., without violating any access through a reference to it.
\paragraph{Buffering} Buffering is a technique used in variety of systems (filesystem page caching, write buffer in multi-processor architectures, translation-lookaside-buffers for translated addresses etc.) to delay the costly operation (committing file pages, flushing buffered writes from write-cache to memory etc.) making updates consistent.
\paragraph{Consistency of Reads of Buffered Writes} Consistency of reads of any value potentially buffered depends on the semantics of the operation making them \emph{observable}. A typical architecture such as Intel-x86 considers the \emph{flush} of buffered writes \emph{non-deterministically} -- i.e., at any time at any number of buffered writes. On the other hand, a typical filesystem can follow a certain policy to commit the in-memory file pages to the disk.   
\subsection{Conclusion \& Contributions}
